---
title: "DataBaseball"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

# On deck: Import the packages we'll need

"There's a package for that!"

```{r}
library(Lahman) # Our source data
library(tidyverse)  # "Metapackage," don't leave home without it!
library(psych)  # Descriptive stats
library(skimr)  # Data profiling
library(writexl) # Write to Excel -- data dump
library(openxlsx) # Write to Excel -- plus 
```

# 1st base: Import the data

We can pull this data from the `Teams` and `Salary` tables of `Lahman`

```{r}
head(Teams)
head(Salaries)
```

Drop some of these Teams columns


```{r}
teams <- Teams %>% 
  select(yearID, teamID, name, attendance, W)

head(teams)
```



We would like to pull the payroll data from `Salaries` and merge it to the data in `Teams`. 

`Salaries` is at the *player* level so we will *aggregate* it, then merge. _Think `VLOOKUP()`!_

```{r}

# Get the total annual salaries by team
payroll <- Salaries %>% 
  group_by(yearID, teamID) %>% 
  summarise(payroll = sum(salary))

head(payroll)

```


Now let's merge this into our `Teams` table:

```{r}

min(Teams$yearID)
min(payroll$yearID)


# Drop the years that don't have payroll data
teams_merged <- teams %>% 
  inner_join(payroll)


head(teams_merged)

```



Sanity check -- does the total payroll equal sum of salaries?


```{r}

sum(teams_merged$payroll) == sum(Salaries$salary)

```

# 2nd base: Profile the data

Lots of ways to profile the data... 

```{r}

# Shape
dim(teams_merged)

# Summary statistics
summary(teams_merged)

```

Many packages exist to aid in the exploration:

```{r}
skim(teams_merged)
```


```{r}
describe(teams_merged)
```


## Shortstop: Data visualization?

This is an easy step to overlook, but it is _crucial_! It leads very well into modeling. 


```{r}
# Distribution of payrolls
ggplot(data = teams_merged, aes(x = payroll)) +
  geom_histogram()

# What could be causing this?
```




```{r}
# Distribution of wins
ggplot(data = teams_merged, aes(x = W)) +
  geom_histogram()
```




# 3rd base: Analyze the data

Depending how statistically informed you want to get, R can "probably" handle it! Let's see what our payroll report might have in store... 


```{r}
# Relationship between payroll and attendance
ggplot(data = teams_merged, aes(x = payroll, y = attendance)) +
  geom_point()
```


```{r}
# Relationship between payroll and wins
ggplot(data = teams_merged, aes(x = payroll, y = W)) +
  geom_point()
```

```{r}
# What if we take the log of payrolls?
teams_merged$log_payroll <- log(teams_merged$payroll)
head(teams_merged)
```


```{r}
# Relationship between payroll and attendance
ggplot(data = teams_merged, aes(x = log_payroll, y = attendance)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

You'd want to back this up with more diagnostics, but you get the idea... 



# Home: Export to Excel

There's definitely more that could be done with our payroll/attendnance analysis, but let's take our victory lap now :) 


If you want a pure data dump to Excel, `writexl` is a good option:


```{r}
# What do you want to write, and to where?
write_xlsx(teams_merged, 'output/writexl.xlsx')
```


```{r}
# What can this function do and not?
# Awesome Easter Egg
?write_xlsx
```

If you would like to fully automate the production from R, `openxlsx` is a good option: 

```{r}
# Add workbook
wb <- createWorkbook()

# Add worksheet of raw data
data_sheet <- addWorksheet(wb, sheetName = 'data')

# Write the dataset as a Table to Excel
writeDataTable(wb, sheet = data_sheet, x = teams_merged, rowNames = FALSE)


# Send plot to the workbook
ggplot(data = teams_merged, aes(x = payroll, y = attendance)) +
  geom_point() + 
  ggtitle('Payroll vs attendance')

# Add the image to the worksheet, set the placement
insertPlot(wb, sheet = data_sheet, startRow = 1, startCol = "I", width = 6, height = 4)

file = 'output/openxlsx.xlsx'

# save workbook to working directory
saveWorkbook(wb = wb, file = file, overwrite = TRUE)
```

We could even add cell formats, conditional formatting and so forth to Excel from R using this package. **Home run!**

Back to the slides for a conclusion... 